---
title: "Exercise week 4"
author: "Nicolas Perez"
date: "November 2017"
output: html_document
---

## Sampling from a normal distribution

In the following simulation, we sample a total of 100 samples from a standard normal distribution and 80 observations per sample. We define $X_i$ as the i-th sample (i=1,2...,100) of the random variable $X$ 

$$X_i \sim \mathcal{N}(0,\,1)$$

```{r, echo=TRUE}
set.seed(116)
observations<-50
samples<-100
datasets<-replicate(samples,rnorm(observations,mean=0,sd=1))
```

## Summarizing the results

For every sampled $X_i$, we compute: 1) absolute bias, 2) standard error, 3) 95% CI Lower Bound and 4) 95% CI Upper bound.

For computing the measures of interest we use the following formulas

$$ bias=|\bar{x}-\mu|$$
Since the population mean is $0$, then the absolute bias can be measured by the absolute value of the mean value of the $i$-th sample.

$$ bias=|\bar{x}|$$
For the standard error we use the estimated standard deviation of the sample and we adjust by the square root of the number of observations

$$ S.E=\hat{\sigma} / \sqrt{n}$$
Finally for the 95% Confidence Intervals, we use the 2.5% and 97.5% quantiles of the normal distribution (it can be shown that the sample mean follows such a distribution but is not part of this exercise) and we build the CI as:

$$ CI_{95\%}= [\bar{x} - 1.96*S.E, \bar{x} + 1.96*S.E] $$

```{r, echo=TRUE}
results<-matrix(NA,nrow=5,ncol=samples)
results[1,]<-abs(colMeans(datasets)-0)
results[2,]<-apply(X=datasets,MARGIN=2,FUN=sd)/sqrt(observations)
results[3,]<-colMeans(datasets)+results[2,]*-1.96
results[4,]<-colMeans(datasets)+results[2,]*1.96
results[5,]<-ifelse((results[4,]<0 | results[3,]>0),1,0)
```

## Including Plots

We can make a plot of the Confidence Intervals containing the true population value for the mean, and given the definition of coverage probabilities for the CIs by Neyman, approximately we should encounter 95% of the CIs to contain the population value for the mean ($0$). 

```{r, echo=TRUE}
plot(c(1:samples),results[3,],type="l", lty=3, ylim=c(-0.6,0.6), main="Coverage Graph, (95% CI)", 
     xlab="sample", ylab="Value")
lines(c(1:samples),results[4,],type="l", lty=3)
abline(h=0, col="blue")
lines(which(results[5,]==1),colMeans(datasets[,results[5,]==1]), type="p", pch=16, col="red")
text(x=1:samples, y=0.07+colMeans(datasets), labels=ifelse(results[5,]==1,c(1:samples),""), col="red") 
```

To create a table we only need to subsample the datasets that are plotted in the graph (already identified through a rule of being or not in the CIs when the results table was created).

```{r, echo=TRUE}
library(knitr)
dataNC<-data.frame(datasets[,results[5,]==1])
colnames(dataNC)<-paste("Sample ", which(results[5,]==1))
kable(dataNC, caption = "Simulated samples that don't cover the population mean")
```